\section{Introduction}

Agentic software systems are increasingly used to execute software tasks, invoke tools, consume budgets, and produce outputs that matter beyond the runtime boundary. As these systems move from laboratory demos to operational workflows, the engineering question is no longer only whether an agent can complete a task. It is also whether the resulting run can be independently reviewed, replayed, and trusted by a third party.

Current practice usually answers this need with observability artifacts such as logs, callback streams, trace records, and framework-specific execution histories. These artifacts are valuable for debugging and operational visibility, but they are not the same as execution evidence. A trace can show that something happened without making the requested action explicit, without preserving a governance decision, without exposing whether execution integrity was checked, and without exporting a bounded receipt that a reviewer can inspect without reconstructing the full runtime context.

This distinction matters because inspectability is weaker than verifiability. An operator may be able to browse a log and infer what likely happened, but an independent reviewer still lacks a stable artifact contract for answering four basic questions: was intent captured, was policy checked, was execution verified, and was a receipt exported? When those questions are left implicit, a run can be observable without being verifiable.

We refer to this mismatch as the \emph{execution evidence gap}. The gap appears whenever a run claims that a task was requested, authorized, executed, and completed, yet the exported artifacts are insufficient to confirm those claims in a structured and reproducible way. In agent stacks this gap is easy to miss because observability tooling is often rich, while audit-oriented evidence surfaces remain ad hoc.

This paper presents an execution evidence architecture for agentic software systems. The proposal is not a new orchestration framework. Instead, it is a composable systems layer that links five stages of runtime semantics: persona projection, intent/action/result objects, governance checkpoints, execution-integrity validation, and bounded audit receipts. The purpose of the architecture is to make agent runs exportable as reviewable evidence bundles.

We instantiate the architecture as a reference path in an artifact package omitted here for double-blind review. The checked-in implementation includes a deterministic 16-task suite, a unified five-file export contract, a third-party review script, a trace-oriented baseline, a framework-shaped external baseline, four ablation modes, and a blinded human-review kit. The resulting harness allows the paper to evaluate execution evidence as an implemented systems concern rather than as a purely conceptual abstraction.

The evaluation compares seven conditions: a local trace-oriented baseline, an external framework-shaped observability approximation, four evidence-chain ablations, and the full evidence-chain configuration. Across these settings we measure five dimensions: explicitness, replayability, tamper sensitivity, audit boundedness, and evidence-stage exposure. The checked-in summaries indicate that the full evidence chain improves every metric over the trace-oriented baseline, while the external baseline primarily improves replayability without recovering the missing evidence semantics. The ablation results further show targeted degradation when individual stages are removed. The architecture contains five stages, but the empirical analysis directly validates the latter four evidence-preserving stages while treating persona projection as an architectural precondition.

The contributions of this paper are threefold:

\begin{itemize}
  \item It formulates execution evidence as a software systems problem that is distinct from runtime observability.
  \item It defines a five-stage execution evidence architecture, while directly evaluating the latter four evidence-preserving stages as an operational review contract.
  \item It implements and evaluates a reproducible reference path showing that evidence-oriented exports preserve review questions that trace-oriented exports do not explicitly retain.
\end{itemize}

The remainder of the paper is organized as follows. Section~\ref{sec:method} introduces the method and reference architecture. Section~\ref{sec:evaluation} describes the evaluation setup. Section~\ref{sec:results} reports the results across baselines and ablations. Section~\ref{sec:related-work} positions the work against adjacent literature. Section~\ref{sec:discussion} discusses limitations, boundary conditions, and implications for software engineering practice.
