\section{Related Work}

This work sits at the intersection of runtime observability, provenance and audit modeling, policy enforcement, and agent-framework execution tooling. The paper does not attempt to replace those lines of work. Instead, it argues that agentic software systems need an additional bridge layer that turns runtime traces into bounded evidence bundles for independent review.

Runtime observability systems standardize logs, metrics, and traces as the primary surface for inspection. OpenTelemetry is representative of this design center \cite{opentelemetry}, while framework-specific tooling such as LangSmith observability and CrewAI tracing exposes agent traces, callbacks, and task timelines for debugging and operations \cite{langsmith-observability,crewai-tracing}. These systems make runs inspectable, but they do not by default provide a review contract that explicitly preserves intent capture, policy persistence, integrity status, and receipt export.

Provenance research emphasizes lineage, responsibility, and reproducibility across heterogeneous systems. PROV-O remains a foundational provenance vocabulary for interoperable provenance interchange \cite{prov-o}, and recent work such as PROV-AGENT extends this tradition to agentic workflows by capturing prompts, responses, and decisions as workflow provenance \cite{prov-agent}. Our execution-evidence layer is narrower than full provenance graphs: it compresses the run into a bounded bundle designed specifically for third-party review questions.

Policy systems show that authorization can be externalized and made explicit. Open Policy Agent is a widely used example of policy-as-code infrastructure \cite{opa}. Recent receipt-oriented work also argues that lightweight execution receipts can offer a practical middle ground between unstructured logs and heavyweight cryptographic proof systems for interactive agents \cite{tool-receipts}. The present paper adopts that general intuition, but evaluates it at the level of exported run bundles, review logic, and controlled architectural ablations.

Relative to these strands, the claim of this paper is modest. Observability surfaces, provenance models, and policy engines each solve part of the problem, but they do not automatically yield an evidence bundle that a third party can inspect without reconstructing the original runtime context. The proposed architecture is intended as that missing bridge, not as a replacement for tracing, provenance, or policy tooling.
