\section{Results}
\label{sec:results}

\subsection{Baseline versus full evidence chain}

Table~\ref{tab:main-comparison} reports the main baseline comparison. The full evidence chain improves every measured dimension over the trace-oriented baseline. Intent capture, policy checking, and receipt export all move from zero observed positive reviews in the baseline to 16 of 16 tasks in the evidence chain. The repository metric records \emph{execution verified} as 11 of 16, but that count should be read in two layers: among the 13 runs that reached the execution stage, 11 completed with verification success and 2 ended as explicit tamper-detected outcomes; the remaining 3 tasks are reviewable policy-blocked outcomes rather than missing bundles.

\input{tables/main_comparison}

The metric changes are large. Average explicitness rises from 1.0 to 5.0, replayability from 3.0 to 5.0, tamper sensitivity from 0.0 to 4.81, audit boundedness from 2.0 to 5.0, and evidence-stage exposure from 1.0 to 5.0. This pattern supports the paper's narrower claim: within a stable exported-bundle contract, the evidence-chain surface preserves review questions that the trace-oriented surface does not explicitly retain.

\subsection{External baseline comparison}

Table~\ref{tab:external-comparison} compares the minimal live CrewAI baseline with the full evidence chain. The external baseline improves replayability relative to the local baseline, reaching an average score of 4.0, because a real CrewAI runtime now produces the task execution surface that is later normalized into the five-file bundle. However, it still leaves explicit intent capture, policy persistence, integrity verification, and receipt export absent. As a result, explicitness, tamper sensitivity, audit boundedness, and evidence-stage exposure remain at 1.0, 0.0, 2.0, and 1.0 respectively.

\input{tables/external_comparison}

This comparison only partially addresses the straw-baseline concern. Because the baseline is a minimal live integration rather than a full production benchmark, the result should be read as evidence that a framework-native default logging surface still leaves major evidence gaps by default, not as a definitive benchmark against every production framework.

\subsection{Ablation results}

Table~\ref{tab:ablation-comparison} summarizes the seven-mode comparison. The ablations show targeted degradation rather than uniform collapse.

\input{tables/ablation_summary}

Removing intent capture lowers explicitness from 5.0 to 3.0 and evidence-stage exposure from 5.0 to 4.0. Under the strengthened review contract it also clears all positive receipt-export verdicts, because the bounded receipt can no longer answer all four review questions once the intent surface is removed. Removing governance lowers explicitness to 2.0, eliminates all positive policy-checked reviews, and changes the task suite behavior by allowing formerly blocked tasks to complete, while tamper sensitivity remains high because integrity checks are still present. Positive receipt-review verdicts also drop to 0 of 16 in that mode, because the bundle can no longer substantiate what was authorized. Removing integrity leaves explicitness and audit boundedness at evidence-chain levels, but drives tamper sensitivity and execution verification to 0.0 and 0 of 16. Removing receipt export preserves most other capabilities but drops audit boundedness from 5.0 to 2.0 and receipt-exported reviews from 16 of 16 to 0 of 16.

Taken together, the ablations indicate that each stage contributes a distinct review function. The evidence chain therefore behaves as a layered architecture rather than as a single monolithic switch.

\subsection{Negative-control falsification checks}

Table~\ref{tab:falsification-summary} reports the new negative-control checks derived from the 11 evidence-chain bundles that originally reviewed as \emph{pass}. Every derived bundle is downgraded from \emph{pass} to \emph{partial}, yielding a 1.0 detection rate across all five falsification scenarios.

\input{tables/falsification_summary}

The pattern of failure is selective rather than indiscriminate. Removing the intent object clears only the intent verdict while leaving policy, execution verification, and receipt export positive in all 11 bundles. Removing the persisted policy decision clears only the policy verdict. Forging a receipt digest clears only the receipt verdict. Tampering with the result payload after export breaks execution verification and receipt verification together because the recomputed digests no longer match the stored bundle. Declaring a tamper outcome without a corroborating digest mismatch is also rejected in all 11 cases, showing that the review script does not merely trust a \texttt{tamper\_detected} label.

These checks do not solve the external-benchmark problem, but they do tighten one internal-validity concern. The repository is no longer only scoring architecture-aligned positives against architecture-aligned metrics; it also demonstrates that the same review contract rejects malformed or over-claimed bundles derived from previously valid runs.

\subsection{Appendix-only reviewer note}

The artifact package includes a blinded reviewer workflow, but its checked-in outputs are synthetic smoke-test artifacts rather than human-subject results. We therefore exclude it from the main result claims and retain it only as appendix-level protocol scaffolding; the details appear in Appendix~\ref{sec:appendix-human-review}.
